---
title: "Make a Plot Assignment"
author: "Grace Walters"
format: pdf
editor: visual
whitespace: small
---

# Make a Plot!

installing ggplot (& other packages I needed to render the PDF)...

```{r}
install.packages("rmarkdown")
```

```{r}
install.packages("ggplot2")
install.packages("knitr")
```

opening the package...

```{r}
library(ggplot2)
```

For my plot, I wanted to visualize how the etymological makeup of words asked in the Scripps National Spelling Bee semifinals & finals has changed from 2013 to 2024 (for instance, what proportion of the words asked in the bee come from Latin and Greek parts? What about French loans?). I work as a spelling bee coach, so trends in this data help me know what's most important to focus on with my students as we prepare for this year's bee :)

my CSV file with my data from my personal computer (I had to do this assignment in my local R because I could not figure out how to get my data to open in a codespace)...

```{r}
etym_props <- read.csv("C:\\Users\\me\\Documents\\National_Spelling_Bee_Etymology_Proportions.csv")

etym_props
```

making the plot...

```{r}
etym_props |>
  ggplot(
  aes(
    x = Year,
    y = Proportion
  )
) +
 geom_line(
    aes(
      color = Language
    )
  ) +
  geom_point(
    aes(
      color = Language
    )
  )+
  scale_x_binned()+
  labs(
    x = "Competition Year",
    y = "% of Total Semis & Finals Words",
    title = "Scripps National Spelling Bee Etymological Makeup Through the Years",
    subtitle = "How the competition has changed over time"
  )+
  theme(
    aspect.ratio = 0.5
  )
```

Looks good for the most part! Some kinks I couldn't iron out:

-   why won't my x axis display my first and last values in "year"?

-   why are my data points plotting between years rather than on them?
